\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-01-26 draft.tex
add footnote on why labor income not included in welfare

Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$. I assume the aggregate labor supply is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade. The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply. I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$. The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$. It is assumed that the production of goods $X$ and $Y$ requires, in addition to labor, the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.

@@

For the case of trade policy, I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.

>>>>

Good $N$ is produced with labor alone with constant returns to scale technology so that $Q_N = l_N$. I assume the aggregate labor supply $l$ is large enough to ensure that the output of good $N$ is large enough to guarantee balanced trade. The supply functions for good $X$ are $Q_X(P_X)$ and $Q_X^*(P_X^*)$ and are assumed strictly increasing and twice continuously differentiable for all prices that elicit positive supply. I also assume $Q_X^*(P_X) > Q_X(P_X)$ for any such $P_X$ so that the home country is a net importer of good $X$. The production structure for good $Y$ is taken to be symmetric, with both demand and supply such that the economy is separable in goods $X$ and $Y$. It is assumed that the production of goods $X$ and $Y$ requires, in addition to labor, the possession of a sector-specific factor that is available in inelastic supply and is non-tradable so that the income of owners of the specific factors is tied to the price of the good in whose production their factor is used.

@@

For the case of trade policy, I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{Labor income $l$ could also be included in both median legislator and executive welfare. This alters none of the results and so would only serve to complicate the exposition.}
